\vskip-0.6in Endogenous Politics and the Design of Trade Institutions

### 2015-01-07 ex_ante.tex
convert lines 46-50 ex_ante to text

\vskip.5in
For simplicity, adding ex-ante lobbying in case of endogenous lobbying ONLY

BS2005 set up that this paper follows holds that governments set TA so that TA will be at optimal level ex-post: gov't will be able to set the applied tariff that matches the political pressure it experiences at the time, and it tunes the TA so that it's exactly the amount of pressure it wants
\begin{itemize}
	\item There is no possibility of ex-ante lobbying in BS2005 framework, so have to figure out what it means
	\item What welfare does each party get at TA stage?
		\begin{itemize}
			\item For lobby is very clear: lobby just knows that this will bind its future action: so paying earlier only matters insofar as it improves ex-post profits
			\item For govts: compare to BS2005, which is all forward looking. With ex-ante, government might get/lose political capital from different interested parties if lobbying is strong enough to outweigh future concerns; otherwise, it's all forward looking---what will future gains and losses be
		\end{itemize}

\newpage
	\item Thus the weight the gov't places on the lobbying industry's profits will change if ex-ante effort is high enough. The trade agreement tariff cap will be set according to $\ga(e^{EA})$ instead of at $\tau^R_{W,e}$ if that would lead to a higher trade agreement tariff (if $\tau = \frac{4(\ga(e^{EA}-1)}{25-4\ga(e^{EA}} < \tau^R_{W,e}$, then government prefers to set ex-post tariff)
			\begin{itemize}
				\item Recall how $\tau^{TA}$ is determined under ex-post lobbying only. Because the TA acts as a commitment device, the government chooses the tariff level $\tau^R_{W,e}$, and implicitly $e^{EP}=e^R_{W,e}$, to maximize the joint political welfare of the two governments as in Equation~\ref{exp:1}.
		\begin{itemize}
			\item Government DOESN'T have to just set trade agreement tariff given what it expects will happen in the future; it can influence lobby's future behavior through the trade agreement (at least on the high side; it can't encourage extra lobbying, which it would like to do in some cases, e.g. exact tariff commitment)
			\item With ex-post lobbying only, at ex-ante stage gov't has to take into account how ex-post lobbying will affect it; picks $\tau^{TA}$ to max joint political welfare subject to this constraint (sets tariff to produce optimal $\ga$ according to $\tau^N$)
		\end{itemize}
				\item This is where gov't loses some measure of control with ex-ante lobbying: 
					\begin{itemize}
						\item If $e^{EA}$ is low enough, ex-ante lobbying has no impact. Government behaves as in ex-post case and chooses $\tau^R_{W,e}$.
						\item If $e^{EA}$ is above some threshold, $\tau^{TA} >\tau^{TA^*}$ b/c $e^{EA}$ now determines $\tau^{TA}$
						\item Gov't still chooses ``optimal'' TA, but it's constrained optimal. Constrained by current pressure.
					\end{itemize}		
				\end{itemize}
	\item Real question: why does the government set non-zero $\tau^R_{W,e}$?
		\begin{itemize}
			\item Because joint welfare is higher when $e^R_{W,e}$ together with $\tau^R_{W,e}$ than it is when $e=0$ and $\tau=0$. These are fundamentally politically motivated governments.
				\begin{itemize}
					\item But depending on how welfare changes as pressure increases, gov't may be looking ahead and manipulating lobbying incentives as shown in Section~\ref{sec:objfcn}
					\item When ex-ante lobbying is possible, in order to respond to that pressure, govts may have to raise $\tau^{TA}$ above $\tau^R_{W,e}$ (lobby won't ever complain if tariff cap is higher than what it asks for)
					\item That is modeled here with a simple cut-off: if $e^{EA}$ is larger than anticipated $e^{EP}$, gov't responds to $e^{EA}$ instead of using TA to influence $e^{EP}$
				\end{itemize}
			\item We will see that this change to the government's preferences has the impacts one would expect, at least in the case of tariff caps: net profits for the lobby (weakly) increase, and net political utility for the government (weakly) decreases
		\end{itemize}
\end{itemize}

@@

\vskip.5in
Tariff cap result
\begin{itemize}
	\item Here, $e^{EP}$ will also work as in body of paper. The earlier stage is sunk, and applied level will match whatever effort lobby puts forth
		\begin{itemize}
			\item It's just that cap could be higher due to ex-ante lobbying
				\begin{itemize}
					\item Without ex-ante lobbying, government chooses $\tau^R_{W,e}$ that leads to ex-post $e$ that maximizes ex-ante joint welfare
		\[
		  W(\ga(e),\tau^R_{W,e}) + W^*(\tau^R_{W,e})
		\]
		knowing they'll set $\tau^N$ in the third stage (i.e. $\tau^N(\ga(e^R_{W,e})) = \frac{8\ga(e^R_{W,e}) - 5}{68 - 8 \ga(e^R_{W,e})} = \tau^R_{W,e}$).
						\begin{itemize}
							\item $\tau^R_{W,e}$ is the optimal unilateral choice of the home government when the lobby is faced with the constraint of the trade agreement
							\item Lobby would max $\pi(\tau(\ga(e))) -e$; instead sets $e = e^R_{W,e} \Rightarrow \tau^N(\ga(e^R_{W,e})) =\tau^R_{W,e}$
						\end{itemize}
					\item Government actually ends up with a lower $e$ mapping into a lower $\ga$ to produce the trade agreement $\tau^R_{W,e}$ according to $\tau^N$, so has to set $\tau^R_{W,e}$ higher to target preferred $\ga$ according to joint max problem [in this symmetric case, identical to unilateral problem]
			\end{itemize}
		\end{itemize}
	\item Lobby decides $e^{EA}$ given that gov't will set cap $\tau^R_{W,e}$ or in ex-ante stage given $e^{EA}$, whichever is higher
		\begin{itemize}
			\item Start with calculus condition
				\[
				  \pi(\tau^E(e^{EA})) - e^{EA} - e^{EP}(e^{EA})
				\]
				When this FOC is operative, government decides tariff cap according to joint decision problem with $e^{EA}$, but then lobby still has to exert effort up to that cap; $e^{EP}$ is the amount the lobby pays according to $\tau^N$ to achieve $\tau^E(e^{EA})$
					\begin{itemize}
						\item FOC
							\begin{equation}
							  \frac{\partial \pi}{\partial \tau} \frac{\partial \tau^E}{\partial \ga} \frac{\partial \ga}{\partial e^{EA}} = 1 +\frac{\partial e^{EP}}{\partial e^{EA}}
								\label{eq:ea}
							\end{equation}
					\end{itemize}
			\item Then, set to zero if
			  \begin{equation}
				  \pi(\tau^E(e^{EA})) - e^{EA} - e^{EP}(e^{EA}) < \pi(\tau^N(e^R_{W,e})) - e^R_{W,e} 
					\label{ine:ea}
				\end{equation}
			\item Intuition: If lobby's incentives to exert pressure ex-ante are strong enough, they overpower the government's ability to use the TA to directly manipulate lobbying incentives ex-post
				\begin{itemize}
					\item Lobby can force government to set cap higher than the level that would optimize its welfare when there is no ex-ante lobbying
					\item Result: With tariff cap, ex-ante lobbying can only (weakly) reduce gov't political welfare. It can only (weakly) increase net profits for lobby (lobby can set $e^{EA}=0$ if ex-ante lobbying would make it worse off)
					\item i.e. tariff with ex-ante lobbying can only be (weakly) higher than without ex-ante lobbying
				\end{itemize}	
		\end{itemize}	
\end{itemize}

@@

					\item If gov't wants more pressure, it should avoid TA. But then loses TOT internalization
					\item If gov't wants less pressure, TA doubly good

@@

							\item Presumes gov't is free to choose which sectors are part of TA, but this is part of a more complicated game; should be able to get payments to leave them out, then payments for unconstrained protection. But GATT rules...

@@

	\item TA mutes lobbying incentives, even if there is ex-ante lobbying

@@

\newpage
\vskip.2in
Compare to ex-ante lobbying \`{a} la MRC.
\begin{itemize}
	\item Cooperative bargaining soln (maximize joint welfare of both governments and both lobbies) instead of extensive form
		\begin{itemize}
			\item leads to bizarre outcome where government will take into account lobby's profits even if $e^{\text{EA}}=0$.
		\end{itemize}
	\item In MRC main treatment, assume lobby has all the bargaining power
		\begin{itemize}
			\item Implies tariff caps and strong bindings equivalent, even though these split the surplus in very different ways. All surplus goes to lobby
			\item In the part that is only in the NBER working paper, they have
				\begin{itemize}
					\item $z=0$ (perfectly immobile capital)
					\item no ex-ante lobbying
					\item government has all the bargaining power
					\item government sometimes chooses free trade---so there IS an incentive to use the TA as a domestic commitment device
				\end{itemize}
			\end{itemize}
	\item In MRC, no matter how tariff is set (cooperatively or non-cooperatively), contribution is $xp(\ov{t})$
		\begin{itemize}
			\item Not true in GH95
			\item In my set-up, lobby's incentives change when tariff-setting process changes
		\end{itemize}
	\item In MRC, weight on lobby's profits is $1+a$ no matter what. Doesn't vary in lobby's efforts. GH not really microfoundations for flexible model as in Long and Vousden (1991) / Baldwin (1987) model; microfoundations for more restrictive version with fixed weights; then decide $\tau$ to change PS, CS, etc. (p. 481 GH94). Can't get shocks as in this literature
		\begin{itemize}
			\item I still don't know exactly why you never want to reign in lobbying in GH; because it's just a conduit for information, and gov't is fully compensated for loss? But is it?
		\end{itemize}
	\item Bargaining power
		\begin{itemize}
			\item It's important in MRC's story: they show how results change if they change bargaining power, and it's dramatic b/c they take an extreme view that lobbies have all the power
			\item I view bargaining power in this environment as an artificial construct---what's going on isn't really bargaining; bargaining is used as a reduced form. But it's a much more complex process
		\end{itemize}
\end{itemize}

>>>>

For simplicity, here I consider ex-ante lobbying when political pressure is endogenous only, i.e. there is no exogenous, probabilistic influence on $\ga$. Since there is no possibility of ex-ante lobbying in the \Textcite{bs2005} framework this paper follows, it must be determined how ex-ante lobbying can best be integrated into the model.

\Textcite{bs2005} posits that governments negotiate the trade agreement so that they will be able to set the applied tariff that best matches the political pressure they experience ex-post. In the case of endogenous political pressure, I have argued that the governments will tune the trade agreement to deliver the amount of lobbying effort that is in their best interest ex-post. How does this change when lobbies can influence the terms of the trade agreement?

In the case of lobbies, a lobby knows that the trade agreement will bind its future action, so exerting effort at the ex-ante stage only matters insofar as it improves ex-post profits net of both ex-ante and ex-post lobbying effort. For the governments, with ex-post pressure only, the governments are completely forward looking: they set the trade agreement with a view toward maximizing expected joint ex-post welfare. In order for ex-ante lobbying to have an effect, it must be that governments also receive some utility from the setting of the trade agreement itself: perhaps they receive or lose political capital from various interested parties, and can be persuaded to forego future benefits if ex-ante lobbying is strong enough to outweigh those future concerns.

Thus I model the governments ex-ante preferences as ... threshold

	\item Thus the weight the gov't places on the lobbying industry's profits will change if ex-ante effort is high enough. The trade agreement tariff cap will be set according to $\ga(e^{EA})$ instead of at $\tau^R_{W,e}$ if that would lead to a higher trade agreement tariff (if $\tau = \frac{4(\ga(e^{EA}-1)}{25-4\ga(e^{EA}} < \tau^R_{W,e}$, then government prefers to set ex-post tariff)
			\begin{itemize}
				\item Recall how $\tau^{TA}$ is determined under ex-post lobbying only. Because the trade agreement acts as a commitment device, the government chooses the tariff level $\tau^R_{W,e}$, and implicitly $e^{EP}=e^R_{W,e}$, to maximize the joint political welfare of the two governments as in Equation~\ref{exp:1}.
		\begin{itemize}
			\item Government DOESN'T have to just set trade agreement tariff given what it expects will happen in the future; it can influence lobby's future behavior through the trade agreement (at least on the high side; it can't encourage extra lobbying, which it would like to do in some cases, e.g. exact tariff commitment)
			\item With ex-post lobbying only, at ex-ante stage gov't has to take into account how ex-post lobbying will affect it; picks $\tau^{TA}$ to max joint political welfare subject to this constraint (sets tariff to produce optimal $\ga$ according to $\tau^N$)
		\end{itemize}
				\item This is where gov't loses some measure of control with ex-ante lobbying: 
					\begin{itemize}
						\item If $e^{EA}$ is low enough, ex-ante lobbying has no impact. Government behaves as in ex-post case and chooses $\tau^R_{W,e}$.
						\item If $e^{EA}$ is above some threshold, $\tau^{TA} >\tau^{TA^*}$ b/c $e^{EA}$ now determines $\tau^{TA}$
						\item Gov't still chooses ``optimal'' trade agreement, but it's constrained optimal. Constrained by current pressure.
					\end{itemize}		
				\end{itemize}
	\item Real question: why does the government set non-zero $\tau^R_{W,e}$?
		\begin{itemize}
			\item Because joint welfare is higher when $e^R_{W,e}$ together with $\tau^R_{W,e}$ than it is when $e=0$ and $\tau=0$. These are fundamentally politically motivated governments.
				\begin{itemize}
					\item But depending on how welfare changes as pressure increases, gov't may be looking ahead and manipulating lobbying incentives as shown in Section~\ref{sec:objfcn}
					\item When ex-ante lobbying is possible, in order to respond to that pressure, govts may have to raise $\tau^{TA}$ above $\tau^R_{W,e}$ (lobby won't ever complain if tariff cap is higher than what it asks for)
					\item That is modeled here with a simple cut-off: if $e^{EA}$ is larger than anticipated $e^{EP}$, gov't responds to $e^{EA}$ instead of using trade agreementto influence $e^{EP}$
				\end{itemize}
			\item We will see that this change to the government's preferences has the impacts one would expect, at least in the case of tariff caps: net profits for the lobby (weakly) increase, and net political utility for the government (weakly) decreases
		\end{itemize}
\end{itemize}

@@

\vskip.5in
Tariff cap result
\begin{itemize}
	\item Here, $e^{EP}$ will also work as in body of paper. The earlier stage is sunk, and applied level will match whatever effort lobby puts forth
		\begin{itemize}
			\item It's just that cap could be higher due to ex-ante lobbying
				\begin{itemize}
					\item Without ex-ante lobbying, government chooses $\tau^R_{W,e}$ that leads to ex-post $e$ that maximizes ex-ante joint welfare
		\[
		  W(\ga(e),\tau^R_{W,e}) + W^*(\tau^R_{W,e})
		\]
		knowing they'll set $\tau^N$ in the third stage (i.e. $\tau^N(\ga(e^R_{W,e})) = \frac{8\ga(e^R_{W,e}) - 5}{68 - 8 \ga(e^R_{W,e})} = \tau^R_{W,e}$).
						\begin{itemize}
							\item $\tau^R_{W,e}$ is the optimal unilateral choice of the home government when the lobby is faced with the constraint of the trade agreement
							\item Lobby would max $\pi(\tau(\ga(e))) -e$; instead sets $e = e^R_{W,e} \Rightarrow \tau^N(\ga(e^R_{W,e})) =\tau^R_{W,e}$
						\end{itemize}
					\item Government actually ends up with a lower $e$ mapping into a lower $\ga$ to produce the trade agreement $\tau^R_{W,e}$ according to $\tau^N$, so has to set $\tau^R_{W,e}$ higher to target preferred $\ga$ according to joint max problem [in this symmetric case, identical to unilateral problem]
			\end{itemize}
		\end{itemize}
	\item Lobby decides $e^{EA}$ given that gov't will set cap $\tau^R_{W,e}$ or in ex-ante stage given $e^{EA}$, whichever is higher
		\begin{itemize}
			\item Start with calculus condition
				\[
				  \pi(\tau^E(e^{EA})) - e^{EA} - e^{EP}(e^{EA})
				\]
				When this FOC is operative, government decides tariff cap according to joint decision problem with $e^{EA}$, but then lobby still has to exert effort up to that cap; $e^{EP}$ is the amount the lobby pays according to $\tau^N$ to achieve $\tau^E(e^{EA})$
					\begin{itemize}
						\item FOC
							\begin{equation}
							  \frac{\partial \pi}{\partial \tau} \frac{\partial \tau^E}{\partial \ga} \frac{\partial \ga}{\partial e^{EA}} = 1 +\frac{\partial e^{EP}}{\partial e^{EA}}
								\label{eq:ea}
							\end{equation}
					\end{itemize}
			\item Then, set to zero if
			  \begin{equation}
				  \pi(\tau^E(e^{EA})) - e^{EA} - e^{EP}(e^{EA}) < \pi(\tau^N(e^R_{W,e})) - e^R_{W,e} 
					\label{ine:ea}
				\end{equation}
			\item Intuition: If lobby's incentives to exert pressure ex-ante are strong enough, they overpower the government's ability to use the trade agreementto directly manipulate lobbying incentives ex-post
				\begin{itemize}
					\item Lobby can force government to set cap higher than the level that would optimize its welfare when there is no ex-ante lobbying
					\item Result: With tariff cap, ex-ante lobbying can only (weakly) reduce gov't political welfare. It can only (weakly) increase net profits for lobby (lobby can set $e^{EA}=0$ if ex-ante lobbying would make it worse off)
					\item i.e. tariff with ex-ante lobbying can only be (weakly) higher than without ex-ante lobbying
				\end{itemize}	
		\end{itemize}	
\end{itemize}

@@

					\item If gov't wants more pressure, it should avoid trade agreement. But then loses TOT internalization
					\item If gov't wants less pressure, trade agreementdoubly good

@@

							\item Presumes gov't is free to choose which sectors are part of trade agreement, but this is part of a more complicated game; should be able to get payments to leave them out, then payments for unconstrained protection. But GATT rules...

@@

	\item trade agreementmutes lobbying incentives, even if there is ex-ante lobbying

@@

\newpage
\vskip.2in
Compare to ex-ante lobbying \`{a} la MRC.
\begin{itemize}
	\item Cooperative bargaining soln (maximize joint welfare of both governments and both lobbies) instead of extensive form
		\begin{itemize}
			\item leads to bizarre outcome where government will take into account lobby's profits even if $e^{\text{EA}}=0$.
		\end{itemize}
	\item In MRC main treatment, assume lobby has all the bargaining power
		\begin{itemize}
			\item Implies tariff caps and strong bindings equivalent, even though these split the surplus in very different ways. All surplus goes to lobby
			\item In the part that is only in the NBER working paper, they have
				\begin{itemize}
					\item $z=0$ (perfectly immobile capital)
					\item no ex-ante lobbying
					\item government has all the bargaining power
					\item government sometimes chooses free trade---so there IS an incentive to use the trade agreement as a domestic commitment device
				\end{itemize}
			\end{itemize}
	\item In MRC, no matter how tariff is set (cooperatively or non-cooperatively), contribution is $xp(\ov{t})$
		\begin{itemize}
			\item Not true in GH95
			\item In my set-up, lobby's incentives change when tariff-setting process changes
		\end{itemize}
	\item In MRC, weight on lobby's profits is $1+a$ no matter what. Doesn't vary in lobby's efforts. GH not really microfoundations for flexible model as in Long and Vousden (1991) / Baldwin (1987) model; microfoundations for more restrictive version with fixed weights; then decide $\tau$ to change PS, CS, etc. (p. 481 GH94). Can't get shocks as in this literature
		\begin{itemize}
			\item I still don't know exactly why you never want to reign in lobbying in GH; because it's just a conduit for information, and gov't is fully compensated for loss? But is it?
		\end{itemize}
	\item Bargaining power
		\begin{itemize}
			\item It's important in MRC's story: they show how results change if they change bargaining power, and it's dramatic b/c they take an extreme view that lobbies have all the power
			\item I view bargaining power in this environment as an artificial construct---what's going on isn't really bargaining; bargaining is used as a reduced form. But it's a much more complex process
		\end{itemize}
\end{itemize}
